---
title: "Untitled"
author: "Gesa Amelung"
date: "7 November 2017"
output:
  html_document: default
  pdf_document: default
---

## Setup

Load the following packages and load the data base access scipt:

```{r setup, echo=TRUE, message=FALSE}
## load packages
require(RPostgreSQL)
require(sf)
require(knitr)

## load access data
path = 'D:/Gesa/Dokumente/Ecotoxicology/RPC'
source(file.path(path, 'amelung_access.R'))

```

## Loading data

Query PostgresSQL data base: bfg_monitoring
Cannot load full psm_samples so far, maybe internet is too weak.. Set LIMIT 100

```{r load-PSM-data, echo=TRUE, message=FALSE}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, dbname = DBname, user = DBuser, host = DBhost, port = DBport, password = DBpassword)

psm_sites = dbGetQuery(con, "SELECT *
                             FROM views.psm_sites_2005
                       ")
psm_samples2 = dbGetQuery(con, "SELECT *
                               FROM views.psm_samples_2005 sam
                               JOIN phch.phch_variables var 
                              ON var.variable_id = sam.variable_id
                         WHERE var.psm_type = 'herbicide'
                         ORDER BY sam.sample_id
                         LIMIT 1000")


psm = dbGetQuery(con, "SELECT *
                               FROM views.psm_samples_2005 sam
                               
                         ORDER BY sam.sample_id
                         LIMIT 1000")

dbDisconnect(con)
dbUnloadDriver(drv)
```

```{r load-diatom-data, echo=TRUE, message=FALSE}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, dbname = DBname, user = DBuser, host = DBhost, port = DBport, password = DBpassword)

require(RPostgreSQL)
require(sf)
require(mapview)

dia_sites = dbGetQuery(con, "SELECT * FROM dia.dia_sites")
dia_sites_geo = st_read_db(con, query = "SELECT * FROM dia.dia_sites LIMIT 10")
dia_sam = dbGetQuery(con, "SELECT *
                       FROM dia.dia_samples
                       WHERE date > '2004-12-31'
                       ")
mapview(dia_sites_geo)

dbDisconnect(con)
dbUnloadDriver(drv)
```

Load physical-chemical data
```{r load-phch-data, echo=TRUE, message=FALSE}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, dbname = DBname, user = DBuser, host = DBhost, port = DBport, password = DBpassword)

# there are more sites than psm_sites (including only pesticides) though they aren't necessary for us.
phch_samples = dbGetQuery(con, "SELECT *
                                FROM phch.phch_samples sam
                                JOIN phch.phch_variables var ON var.variable_id = sam.variable_id
                                WHERE var.name IN ('Wassertemperatur', 'Sulfat');")

# 596477 observations

dbDisconnect(con)
dbUnloadDriver(drv)
```

## Question: How to compare diatom and psm sampling sites?
Where were samples taken at the same location?

Which column should I use to compare both datasets?
Random check if site_id exists in dia_sam.
Result: site_id is not the same in psm_samples and dia_sam.
Maybe labelling system is still the same, samples were just not always taken at same location?
```{r echo=TRUE, message=FALSE}
# Random site_id taken from psm_samples
dia_sam[dia_sam$site_id=="BW_10001",] #no result
dia_sam[dia_sam$site_id=="RP_10002",] #no results
# Random site_id taken from dia_samples as a cross-check if code is correct.
dia_sam[dia_sam$site_id=="BW_AI025.00",] #185 results

```
Is site_name, site_id and site_nr the same in psm_sites dataset? 
Find out by counting distinct occurence of them.
Result: site_id and site_name does not mean the same.
Site_id and site_nr might be approximately the same.
``` {r echo=TRUE, message=FALSE}
# Complete data set has to be queried before!!
psm_siteid_list <- unique(psm_sites$site_id)
nrow(as.data.frame(psm_siteid_list)) # 3116 different entries for site_id in psm_sites

psm_sitename_list <- unique(psm_sites$site_name)
nrow(as.data.frame(psm_sitename_list)) # 2746 different entries for site_name in psm_sites

psm_sitenr_list <- unique(psm_sites$site_nr)
nrow(as.data.frame(psm_sitenr_list)) # 3114 different entries for site_nr in psm_sites

```

Random check if site_nr in dia_sites and psm_sites are the same.
Result: site_nr in psm_sites and dia_sites are not the same.
Maybe labelling system is still the same, samples were just not always taken at same location?
Check the other way around. Are site_nr of dia_sites found in psm_sites?
Result: Positive. Dia samples were probably not taken at all psm sampling locations.
```{r echo=TRUE, message=FALSE}
# Random site_nr taken from psm_sites
dia_sites[dia_sites$site_nr=="CAC029",] # no results
dia_sites[dia_sites$site_nr=="21094",] # no results
dia_sites[dia_sites$site_nr=="CEN901",] # no results
# Random site_nr taken from dia_sites as a cross-check to see if code is correct.
dia_sites[dia_sites$site_nr=="CEZ016",] # 1 result

# Check the other way around. Are site_nr of dia_sites found in psm_sites? YES.
psm_sites[psm_sites$site_nr=="CEZ016",] # 1 result
dia_sites[dia_sites$site_nr=="CEZ016",] # 1 result
psm_sites[psm_sites$site_nr=="CKO317",] # 1 result
dia_sites[dia_sites$site_nr=="CKO317",] # 1 result
psm_sites[psm_sites$site_nr=="21081",] # 1 result
dia_sites[dia_sites$site_nr=="21081",] # no results
psm_sites[psm_sites$site_nr=="105703",] # 1 result
dia_sites[dia_sites$site_nr=="105703",] # 1 result


```

Random check if site_id in dia_sites and psm_sites are the same.
Result: NO. Only sometimes it works.
``` {r echo=TRUE, message=FALSE}
psm_sites[psm_sites$site_id=="BW_CEZ016",] # 1 result
dia_sites[dia_sites$site_id=="BW_CEZ016",] # no result
psm_sites[psm_sites$site_id=="BW_CKO317",] # 1 result
dia_sites[dia_sites$site_id=="BW_CKO317",] # no result
psm_sites[psm_sites$site_id=="BY_3390",] # 1 result
dia_sites[dia_sites$site_id=="BY_3390",] # 1 result


```
# Result
site_id cannot be used to connect psm_sites and dia_sites. Instead site_nr must be used
since attribut is found in both tables.


## Compare both sample datasets
Filter sites where diatom and herbicide data was measured in the same year.
``` {r compare-sample-datasets, echo=TRUE, message=FALSE}
# Join both sample data frames. Add a column identifying which source each row comes from.
dia_sam$table <- "dia"
psm_samples$table <- "psm"
dia_sam_bind <- dia_sam[,c("date", "site_id", "table")]
psm_sam_bind <- psm_samples[,c("date", "site_id", "table")]
samples_bind <- rbind(dia_sam_bind,psm_sam_bind) # combine both sample tables

```

Create data frames with samples only from 2005.
``` {r create-year-column, echo=TRUE, message=FALSE}
# Create year column from date column
dia_sam$year <- as.numeric(format(dia_sam$date,'%Y'))
psm_samples$year <- as.numeric(format(psm_samples$date,'%Y'))

# Create data frames with samples only from 2005
dia_sam_05 <- dia_sam[dia_sam$year=="2005",]
psm_sam_05 <- psm_samples[psm_samples$year=="2005",]

# common <- intersect(dia_sam_05[,"site_id"], psm_sam_05[,"site_id"]) # results in empty data

```
# Want to compare tables on year and site_id:
Extract year from date.
```{r  extract-year-from-date echo=TRUE, message=FALSE}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, dbname = DBname, user = DBuser, host = DBhost, port = DBport, password = DBpassword)

sam_tab = dbGetQuery(con, " SELECT dsam.site_id, dsam.date, psam.date, 
                    EXTRACT(YEAR FROM dsam.date) AS sam_year,
                    EXTRACT(YEAR FROM psam.date) AS psm_year
                     FROM dia.dia_samples AS dsam
                     INNER JOIN views.psm_samples_2005 AS psam 
                     ON dsam.site_id = psam.site_id
                    LIMIT 100");


dbDisconnect(con)
dbUnloadDriver(drv)
```

Also, should I shorten the species name to include only genus and species and no further description?

Select all diatom samples where environmental variable data was measured at the same site in the same year.
```{r  select-diasamples-join-site-year, echo=TRUE, message=FALSE}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, dbname = DBname, user = DBuser, host = DBhost, port = DBport, password = DBpassword)


# including all years --> works, but veeeeery slowly
join = dbGetQuery(con, " SELECT DISTINCT ON (dsam.site_id, dsam.date, dsam.taxon) dsam.site_id, dsam.date, taxon, psam.date
                   FROM dia.dia_samples AS dsam
                   INNER JOIN views.psm_samples_2005 AS psam
                   ON dsam.site_id = psam.site_id
                   WHERE (dsam.date BETWEEN '2005-01-01' AND '2005-12-31'
                   AND psam.date BETWEEN '2005-01-01' AND '2005-12-31')
                OR (dsam.date BETWEEN '2006-01-01' AND '2006-12-31'
                   AND psam.date BETWEEN '2006-01-01' AND '2006-12-31')
                OR (dsam.date BETWEEN '2007-01-01' AND '2007-12-31'
                   AND psam.date BETWEEN '2007-01-01' AND '2007-12-31')
                OR (dsam.date BETWEEN '2008-01-01' AND '2008-12-31'
                   AND psam.date BETWEEN '2008-01-01' AND '2008-12-31')
               OR (dsam.date BETWEEN '2009-01-01' AND '2009-12-31'
                   AND psam.date BETWEEN '2009-01-01' AND '2009-12-31')
                OR (dsam.date BETWEEN '2010-01-01' AND '2010-12-31'
                   AND psam.date BETWEEN '2010-01-01' AND '2010-12-31')
                 OR (dsam.date BETWEEN '2011-01-01' AND '2011-12-31'
                   AND psam.date BETWEEN '2011-01-01' AND '2011-12-31')
                 OR (dsam.date BETWEEN '2012-01-01' AND '2012-12-31'
                   AND psam.date BETWEEN '2012-01-01' AND '2012-12-31')
                 OR (dsam.date BETWEEN '2013-01-01' AND '2013-12-31'
                   AND psam.date BETWEEN '2013-01-01' AND '2013-12-31')
                OR (dsam.date BETWEEN '2014-01-01' AND '2014-12-31'
                   AND psam.date BETWEEN '2014-01-01' AND '2014-12-31')
                OR (dsam.date BETWEEN '2015-01-01' AND '2015-12-31'
                   AND psam.date BETWEEN '2015-01-01' AND '2015-12-31')
                   ORDER BY dsam.date, dsam.site_id, dsam.taxon
                   LIMIT 100000")
# results in 47127 observations

# Include to select for sites where relevant variables were measured.
#JOIN phch.phch_variables AS var 
#                      ON var.variable_id = psam.variable_id
#                    WHERE var.psm_type = 'herbicide'

dbDisconnect(con)
dbUnloadDriver(drv)
```
```{r  join-relevant-variables, echo=TRUE, message=FALSE}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, dbname = DBname, user = DBuser, host = DBhost, port = DBport, password = DBpassword)

join2 = dbGetQuery(con, " SELECT DISTINCT ON (dsam.site_id, dsam.date, dsam.taxon) dsam.site_id, dsam.date, taxon, iz_n, psam.date, var.psm_type, var.name
                   FROM dia.dia_samples AS dsam
                   INNER JOIN views.psm_samples_2005 AS psam
                   ON dsam.site_id = psam.site_id
                   INNER JOIN phch.phch_variables var 
                   ON var.variable_id = psam.variable_id 
                     WHERE var.psm_type = 'herbicide'
                   AND ((dsam.date BETWEEN '2005-01-01' AND '2005-12-31'
                   AND psam.date BETWEEN '2005-01-01' AND '2005-12-31')
                OR (dsam.date BETWEEN '2006-01-01' AND '2006-12-31'
                   AND psam.date BETWEEN '2006-01-01' AND '2006-12-31')
                OR (dsam.date BETWEEN '2007-01-01' AND '2007-12-31'
                   AND psam.date BETWEEN '2007-01-01' AND '2007-12-31')
                OR (dsam.date BETWEEN '2008-01-01' AND '2008-12-31'
                   AND psam.date BETWEEN '2008-01-01' AND '2008-12-31')
               OR (dsam.date BETWEEN '2009-01-01' AND '2009-12-31'
                   AND psam.date BETWEEN '2009-01-01' AND '2009-12-31')
                OR (dsam.date BETWEEN '2010-01-01' AND '2010-12-31'
                   AND psam.date BETWEEN '2010-01-01' AND '2010-12-31')
                 OR (dsam.date BETWEEN '2011-01-01' AND '2011-12-31'
                   AND psam.date BETWEEN '2011-01-01' AND '2011-12-31')
                 OR (dsam.date BETWEEN '2012-01-01' AND '2012-12-31'
                   AND psam.date BETWEEN '2012-01-01' AND '2012-12-31')
                 OR (dsam.date BETWEEN '2013-01-01' AND '2013-12-31'
                   AND psam.date BETWEEN '2013-01-01' AND '2013-12-31')
                OR (dsam.date BETWEEN '2014-01-01' AND '2014-12-31'
                   AND psam.date BETWEEN '2014-01-01' AND '2014-12-31')
                OR (dsam.date BETWEEN '2015-01-01' AND '2015-12-31'
                   AND psam.date BETWEEN '2015-01-01' AND '2015-12-31'))
                   ORDER BY dsam.date, dsam.site_id, dsam.taxon
                   ")
# results in 45980 observations


dbDisconnect(con)
dbUnloadDriver(drv)
```

```{r  join-2005, echo=TRUE, message=FALSE}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, dbname = DBname, user = DBuser, host = DBhost, port = DBport, password = DBpassword)

# since iz_n (species abundance) is often empty, should I instead choose cover as abundance value?

join2005 = dbGetQuery(con, " SELECT DISTINCT ON (dsam.site_id, dsam.taxon, dsam.date) dsam.site_id, dsam.date, taxon, dsam.iz_n, cover, psam.date, var.psm_type, var.name
                   FROM dia.dia_samples AS dsam
                   INNER JOIN views.psm_samples_2005 AS psam
                   ON dsam.site_id = psam.site_id
                   INNER JOIN phch.phch_variables var 
                   ON var.variable_id = psam.variable_id 
                  WHERE var.psm_type = 'herbicide'
                   AND (dsam.date BETWEEN '2005-01-01' AND '2005-12-31'
                   AND psam.date BETWEEN '2005-01-01' AND '2005-12-31')
                  ORDER BY dsam.date, dsam.site_id, dsam.taxon
                   ")

dbDisconnect(con)
dbUnloadDriver(drv)
```
Transform data with reshape2 package to get from long to wide data for the site_id.
```{r  transform-data-to-wide, echo=TRUE, message=FALSE}


tab05 <- data.frame(join2005$taxon, join2005$site_id, join2005$date,as.numeric(join2005$cover))
names(tab05) <- c("taxon", "site_id", "date", "cover" )
names(tab05)
require(reshape2)
tab05cast <- dcast(tab05, taxon + date ~ site_id, fill = 0)

# Like this, same species appears several times if it was sampled on different sites on
# several dates in the year.


# tab05cast <- dcast(tab05, taxon ~ site_id, fill = 0)
# aggregation function needed if variables do not identify a single observation for
# each output cell. Defaults to length (with a message) if needed but not specified.
# eg Achnanthes minutissima var. minutissima is present twice at site_id NW_137315
# Answer: Result of if same site was sampled twice in one year.

# creates correct type of data but cover data is not included
# dcast ( tab, taxon ~ site_id + year ) would also be possible


```

```{r  select-diasamples-join-site-year, echo=TRUE, message=FALSE}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, dbname = DBname, user = DBuser, host = DBhost, port = DBport, password = DBpassword)

#query takes for ever!
join6 = dbGetQuery(con, "SELECT dsam.site_id, dsam.date AS d_date, taxon
                   FROM dia.dia_samples AS dsam
                   WHERE site_id IN 
                   (
                   SELECT site_id
                   FROM views.psm_samples_2005 AS psam
                   WHERE EXTRACT ('year' FROM psam.date) = 2005
                   AND EXTRACT ('year' FROM dsam.date) = 2005
                     )
                   ")


dbDisconnect(con)
dbUnloadDriver(drv)
```


# Creating a data frame
```{r  create-data-frame, echo=TRUE, message=FALSE}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, dbname = DBname, user = DBuser, host = DBhost, port = DBport, password = DBpassword)



dia_sam_year = dbGetQuery(con, " SELECT *, 
                    EXTRACT(YEAR FROM dsam.date) AS year
                     FROM dia.dia_samples AS dsam
                    WHERE date >= '2005-01-01'
                    ")
dia_sam_year = dbGetQuery(con, " SELECT dsam.id_rs, dsam.id_pn, dsam.date, dsam.dv, dsam.iz_n, dsam.iz_b, dsam.cover, dsam.site_id, dsite.site_nr, 
                    EXTRACT(YEAR FROM dsam.date) AS year
                     FROM dia.dia_samples AS dsam
                     JOIN dia.dia_sites AS dsite 
                     ON dsam.site_id = dsite.site_id                  
                    WHERE date >= '2005-01-01'
                    ")

psm_sam_year = dbGetQuery(con, " SELECT  psam.sample_id, psam.site_id,psam.date, psam.qualifier, psam.value, psam.unit, var.name, var.cas, var.pgroup, var.psm_type, var.pan_class, psite.site_nr,
                    EXTRACT(YEAR FROM psam.date) AS year
                     FROM views.psm_samples_2005 AS psam
                      JOIN phch.phch_variables var 
                      ON var.variable_id = psam.variable_id
                      JOIN views.psm_sites_2005 AS psite 
                      ON psam.site_id = psite.site_id                    
                      WHERE date >= '2005-01-01'
                     AND var.psm_type = 'herbicide'
                    LIMIT 1000")

# Create the data frame
samples <- merge(dia_sam_year, psm_sam_year, by = c("site_nr","year"))
samples <- merge(dia_sam_year, psm_sam_year, by = "year")
samples2 <- merge(dia_sam_year, psm_sam_year, by = "site_nr")
# Works for year but not for site_nr.
#Error: cannot allocate vector of size 4.4 Gb -> if 100000 objects in both dataframes
# merge by site_id results in emtpy data
# results in empty data

dbDisconnect(con)
dbUnloadDriver(drv)
```

## Shorten taxon name
Not finished yet.
```{r  shorten-taxon-name, echo=TRUE, message=FALSE}
taxa_list <-unique(dia_herb_year$taxon)
taxa_cha <- as.character(taxa_list)
species_list <- sort(taxa_cha) #list of all species
species_list
nrow(as.data.frame(species_list))

taxa_genus <- sapply(strsplit(taxa_cha, ' '), '[', 1) # keeps only genus name
taxa_species <- sapply(strsplit(taxa_cha, ' '), '[', 2) # keeps only species name
genus_list<- sort(unique(taxa_temp))    #list of all genera
genus_list
nrow(as.data.frame(genus_list))

# combine genus and species name
taxon <- paste

```
There are 484 different species in the data set if limit is set to 100,000.

## Taxon frequency
Are there any rare species? Marchant (2002) suggests to remove rare species and Cao et al.(2001) suggests this if samples were taken over a large spatial scale which holds true for the present data. So the question is what a suitable definition for rare species is. E.g. species that occur in only 1 sample.

```{r  taxon-frequency, echo=TRUE, message=FALSE}

# taxon_count <- table(dia_sam$dv)
taxon_count <- table(join2$taxon)
# taxon_count <- table(dia_sam$taxon)
summary(taxon_count)
taxon_count <-as.data.frame(taxon_count)
taxon_count <- taxon_count[order(taxon_count$Freq),]
head(taxon_count)
freqn <- nrow(taxon_count) # 931
freq3 <- nrow(taxon_count[taxon_count$Freq<5,]) # 383
freq2 <- nrow(taxon_count[taxon_count$Freq<3,]) # 274
freq1 <- nrow(taxon_count[taxon_count$Freq<2,]) # 172
hist(taxon_count$Freq, breaks = 50)

# relative frequency of species that are only present at 1 site
freq1/freqn*100


# Sollte ich seltene Arten weglassen? Wenn ja, welche Arten stufe ich als selten ein?
```
The distribution of species frequency s very left-skewed. Most speces are present in less than ~200 samples and few are present in more than 200 samples.

```{r  remove-rare-species, echo=TRUE, message=FALSE}
# Removal of the 172 species that occur only at 1 site.
```

```{r  read-taxalisten-data, echo=TRUE, message=FALSE}
samples_taxon <- merge(dia_sam_year, diatom_dat, by.x="dv", by.y="DV")
diatom_dat1<- read.table("C_Proben_Taxalisten.txt",header=T,sep="\t",dec=".")

```

```{r  suggestions-andreas, echo=TRUE, message=FALSE}
require(data.table)    # good for subsampling --> datacamp course
iris_dt <- as.data.table(iris)
class(iris_dt)
head(iris_dt)
iris_dt[i = 1:100, j = mean(Sepal.Length), by = Species]
diatom_dat1 <- fread("C_Proben_Taxalisten.txt") # schnellere funktion als read.csv usq. auch: fwrite()


require(dplyr)     # good for subsampling

iris$Sepal.Length
# with base package:
aggregate(iris$Sepal.Length, by = list(iris$Species), mean)

```


```{r get-variables, echo=TRUE, message=FALSE}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, dbname = DBname, user = DBuser, host = DBhost, port = DBport, password = DBpassword)


var_names = dbGetQuery(con, "SELECT DISTINCT var.name
                                FROM phch.phch_samples sam
                                JOIN phch.phch_variables var ON var.variable_id = sam.variable_id
                                ")
# Nitrit, Nitrat-Stickstoff, Kalium, Sichttiefe, Chlorid, Nitrat, Gesamt-Phosphor, Sauerstoffgehalt, Sauerstoffzehrung nach 5 (7) Tagen ohne Hemmer, Gesamt-Stickstoff, Stickstoff, Wasserstand (OFW), pH-Wert (Feld) Ammonium-Stickstoff, Sauerstoffzehrung nach 7 Tagen mit Hemmer, BSB5, sedimentiert, Sauerstoffzehrung nach 21 Tagen mit Hemmer, Trübung (physiko-chem. Messung), Elektrische Leitfähigkeit (20°C), Nitrit-Stickstoff, Natrium, Sulfat, Ammonium, Abfluss, spektraler Absorptionskoeffizient 436nm, freie Kohlensäure, Sauerstoffsättigung, Stickstoff, Hydrogenkarbonat, Eisen, Wassertemperatur, Magnesium, Gesamtphosphor als PO4, Phosphor, Chemischer Sauerstoffbedarf

dbDisconnect(con)
dbUnloadDriver(drv)
```